{\small} CUBEP3M: Closer to optimal N-body computations

### 2012-04-16 Introduction.tex
Completed first draft of Introduction.tex

For instance, a measurement of the baryonic acoustic oscillation (BAO) dilation scale can provide
tight constraints on the dark energy equation of scale \citep{Seo2003,2005}.
The most optimal estimates of the uncertainty requires the knowledge of the 
matter power spectrum covariance matrix, which is only accurate when measured from a large sample 
of N-body simulations \citep{RH2005, japan2,japan3, Ngan}.
For the same reasons, the most accurate estimates of weak gravitational lensing signal is obtained
by propagating photons in past light cones that are extracted from simulated density fields \citep{HarnoisLudo, Takahashi }.
{\bf (Ilian, could you say something about reionization here?)}

@@

{\small PMFAST} (\cite{PMFAST}, MPT hereafter) is one of the first code that is designed such as to optimize the PM algorithm,
both in terms of speed and memory usage. It uses a two-level mesh algorithm based on the gravity solver of \cite{TracPen2003},
The long range gravitational force is computed on a  grid four times coarser, such as to minimize the communication time
and to fit in system's memory. The short range is computed locally on a finer mesh, which can be as tight as the memory allows.
This enable the code to evolve very large cosmological systems both rapidly and accurately, on relatively modest clusters.

@@

>>>>

For instance, a measurement of the baryonic acoustic oscillation (BAO) dilation scale can provide
tight constraints on the dark energy equation of scale \citep{Seo2003,2005}.
The most optimal estimates of the uncertainty requires the knowledge of the 
matter power spectrum covariance matrix, which is only accurate when measured from a large sample 
of N-body simulations \citep{RH2005, japan2,japan3}.
For the same reasons, the most accurate estimates of weak gravitational lensing signal is obtained
by propagating photons in past light cones that are extracted from simulated density fields \citep{White, Jain, Takahashi}.
{\bf (Ilian, could you say something about reionization here? Just general references as to why one needs N-body codes to perform 
accurate predictions/forecasts, etc.)}

@@

@@

{\small PMFAST} (\cite{PMFAST}, MPT hereafter) is one of the first code that is designed such as to optimize the PM algorithm,
both in terms of speed and memory usage. It uses a two-level mesh algorithm based on the gravity solver of \cite{TracPen2003},
The long range gravitational force is computed on a  grid four times coarser, such as to minimize the communication time
and to fit in system's memory. The short range is computed locally on a finer mesh, and only the local sub-volume needs 
to be stored at a given time, allowing for parallel computation.
This this setup enable the code to evolve very large cosmological systems both rapidly and accurately, on relatively modest clusters.
One of the main advantage of {\small PMFAST} over other PM codes is that the number of large arrays is minimized,
and the global {\small MPI} communications are cut down to the minimum: for passing particle at the beginning of each time step,
and  for computing the long range FFTs.
As described in MPT, access to particles is accelerated with the use of linked lists, deletion of `ghost' particles
in buffer zones is done at the same time as particles are passed to adjacent nodes,
and the global FFTs are performed with a slab decomposition of the volumes via a special file transfer interface, 
designed specifically to preserve a high processor load.

Since its first published version, {\small PMFAST} has evolved in many aspects. 
The first major improvement was to transform the volume decomposition in multi-node configurations 
from slabs to cubes. One of the problem with slabs is that they do not scale well to large runs: 
as the number of cells per dimension increases, the thickness of each slab shrinks rapidly,
until it reaches the hard limit of a single cell layer.  With this enhancement came a new name, {\small CUBEPM},
which soon after incorporated particle-particle (pp) interactions at the sub-grid level, and the code was renamed
{\small CUBEP3M}. It now includes a significant number of new features since MPT : the pp force
can be extended to an arbitrary range, the size of the redshift jump can be constrained for improved accuracy during the first time steps,
a runtime halofinder has been implemented, the expansion has also been generalized to include a redshift dependent equation of state of dark energy, there is a system of unique particle identification which can be switched on or off, and the initial condition generator has been generalized as to include non-Gaussian features
{\bf (anything else?)}.
{\small CUBEP3M} has already been involved in a number of scientific applications over the last few years,
spanning the field of weak lensing \citep{Sanaz_Lu, Lu_dore, Dore_Lu, Harnois_Vafaei, LuCMBLensing},  BAO
 \citep{HarnoisPen, Ngan,Zhang, Yu, HarnoisYu in prep.}, re-ionization \citep{Ilian...}, {\bf (anymore cubep3m papers to cite?)}

This paper aims at presenting and quantifying these new features that make {\small CUBEP3M} one of the most competitive public N-body code.
It is structured as follow: XXXX

### 2012-04-16 paper.tex
Completed first draft of Introduction.tex

\title[{\small} CUBEP3M: Closer to optimal N-body computations]{{\small} CUBEP3M: Closer to optimal N-body computations}
\author[Joachim Harnois-D\'{e}raps, Ue-Li Pen, Hugh Merz, Ilian T. Iliev, Hy Trac]{Joachim Harnois-D\'{e}raps$^{1,2}$ 
\thanks{E-mail: jharno@cita.utoronto.ca},  Ue-Li Pen$^{1}$ \thanks{E-mail: pen@cita.utoronto.ca} 
Hugh Merz$^{3}$ \thanks{E-mail: xxx},  Ilian T. Iliev$^{4}$ \thanks{E-mail: xxx} and
Hy Trac$^{5}$ \thanks{E-mail: xxx}\\
%\footnotemark[1]\thanks{This file has been amended to
%highlight the proper use of \LaTeXe\ code with the class file.
%These changes are for illustrative purposes and do not reflect the
%original paper by A. V. Raveendran.}\\
$^{1}$Canadian Institute for Theoretical Astrophysics, University of
Toronto, M5S 3H8, Canada\\
$^{2}$Department of Physics, University of Toronto, M5S 1A7, Ontario,  Canada\\
$^{3}$Department of XXX\\
$^{4}$Department of XXX\\
$^{5}$Department of XXX}

>>>>

\title[{\small} CUBEP3M: Closer to optimal N-body computations]{{\small} CUBEP3M: Closer to optimal N-body computations}
\author[Joachim Harnois-D\'{e}raps, Ue-Li Pen, Hugh Merz, Ilian T. Iliev]{Joachim Harnois-D\'{e}raps$^{1,2}$ 
\thanks{E-mail: jharno@cita.utoronto.ca},  Ue-Li Pen$^{1}$ \thanks{E-mail: pen@cita.utoronto.ca} 
Hugh Merz$^{3}$ \thanks{E-mail: xxx} and  Ilian T. Iliev$^{4}$ \thanks{E-mail: xxx}\\
%\footnotemark[1]\thanks{This file has been amended to
%highlight the proper use of \LaTeXe\ code with the class file.
%These changes are for illustrative purposes and do not reflect the
%original paper by A. V. Raveendran.}\\
$^{1}$Canadian Institute for Theoretical Astrophysics, University of
Toronto, M5S 3H8, Canada\\
$^{2}$Department of Physics, University of Toronto, M5S 1A7, Ontario,  Canada\\
$^{3}$Department of XXX\\
$^{4}$Department of XXX\\
$^{5}$Department of XXX}
